### 2014-02-01 paper/workspace/main-example-filip.tex
Added version of paper draft with Filip's example

\documentclass{article}

\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{color}
\usepackage{listings}
\usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
\usepackage{graphicx}

\usepackage{tikz}
\usetikzlibrary{arrows,shapes,shapes.multipart,shapes.geometric,
	snakes,automata,backgrounds,petri,calc,positioning}

\newcommand{\bfc}{{\sc Bfc}}
\newcommand{\mist}{{\sc Mist}}
\newcommand{\iic}{{\sc IIC}}
\newcommand{\pnerf}{{\sc Pnerf}}
\newcommand{\zthree}{{\sc Z3}}
\newcommand{\ttt}[1]{\texttt{#1}}

\tikzstyle{place}=[circle,thick,draw=black!75,fill=white!20,
  	minimum size=6mm]
\tikzstyle{red place}=[place,draw=red!75,fill=red!20]
\tikzstyle{blue place}=[place,draw=blue!75,fill=blue!20]
\tikzstyle{transition}=[rectangle,thick,draw=black!75,
  	fill=black!20,minimum size=4mm]

\tikzstyle{state}=[draw, ellipse, aspect=2]
\tikzstyle{action}=[draw, rectangle, align=center]
\tikzstyle{decision}=[draw, diamond, aspect=2, align=center]
\tikzstyle{print}=[draw, trapezium, trapezium left angle=70, trapezium right angle=-70]
\tikzstyle{every edge}=[draw, ->, >=stealth, shorten >=2pt, shorten <=2pt]

\begin{document}

\title{Report on a Safety Checker for Petri nets}
\author{Philipp Meyer \and Rusl\'{a}n Ledesma-Garza}
% \institute{Technische Universit\"at M\"unchen}
\date{Mon Nov 25 14:24:10 CET 2013}

\section{Preliminaries}

\begin{verbatim}
* Petri net:
 N = (S, T, E, M0), where
 S : places
 T : transitions
 E : edges
 M0: initial marking

* Property:
 P, linear arithmetic formula
\end{verbatim}
\iffalse
\begin{verbatim}
 Examples:
  x + y < 0
  x + y < 0 \\and x + z > 0
  x + y < 0 \\or  x + z > 0

 ~P = (x + y <  0 \\and x + z >  0) \\or  x + z >  0
  P = (x + y >= 0 \\or  x + z <= 0) \\and x + z <= 0
\end{verbatim}
\fi

\newpage
\section{Method Safety}

The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
We present the method Safety by example on a modified Petri net from [Esparza1997].

\begin{verbatim}
* Code

TODO

* Property: z is never marked

* Petri net:

\end{verbatim}

\begin{center}
  \includegraphics[width=6cm]{example-filip}
\end{center}

\begin{verbatim}
* Property P: z = 0
\end{verbatim}

\newpage
\begin{verbatim}
* Method Safety:

  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
\end{verbatim}

\begin{center}
  \begin{tikzpicture}
    \node[state] (begin) {BEGIN};
    \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
    \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
    \node[print, right=of satc] (yes) {N satisfies P};
    \node[print, below=of satc] (dontknow) {Don't know};
    \node[state, right=of yes] (end1) {END};
    \node[state, below=of dontknow] (end2) {END};

    \draw (begin) edge (c);
    \draw (c) edge (satc);
    \draw (satc) edge node[above]{NO} (yes);
    \draw (yes) edge (end1);
    \draw (satc) edge node[right]{YES} (dontknow);
    \draw (dontknow) edge (end2);
  \end{tikzpicture}
\end{center}

\begin{verbatim}
* Property of state constraints C: If C U {\neg P} is unsat then N |= P.

* Example:
  - State constraints C:

    Place equations:

    x = 1 - t
    ^   ^   ^
    |   |   |
    |   |   |
    |   |   |
    |   |   |
    |   |   |
    |   |  # of tokens taken from x
    |   |
    |   initial number of tokens in x
    |
    number of tokens in x

    y = 0
    z = 0 + t
    w = 0 + t
\end{verbatim}

\newpage

\begin{verbatim}
    Non-negativity conditions:

    x >= 0
    y >= 0
    z >= 0
    w >= 0
    t >= 0

  - Negated Property \neg P:

    z >= 1

* Place equation:

  For a given place s the place equation is

  # of tokens in s = initial number of tokens of place s
                     + # times each input transition of s fires
                     - # times each output transition of s fires

* Non-negativity conditions:

  # of tokens in place s           >= 0
  # of times transition t is fired >= 0

* Subprocedure \mathcal{C}:

  Input:
    (S, T, E, M0) : Petri net
  Output:
    C : State constraints

  Pseudocode:

  \end{verbatim}
\begin{align*}
  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
  \right) \right) \land \\
  & \left( \bigwedge_{s \in S} s \ge 0 \right) \land
    \left( \bigwedge_{t \in T} t \ge 0 \right)
\end{align*}

\newpage

\section{Method Safety by Refinement}

The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.

\begin{verbatim}
* Method Safety by Refinement:

  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
\end{verbatim}
\begin{center}
\begin{tikzpicture}
  \node[state] (begin) {BEGIN};
  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
                                    	$\delta:=\Delta(A_\theta)$\\
                                    	$C:=C \cup \{\delta\}$};
  \node[print, right=of satc] (yes) {N satisfies P};
  \node[print, right=of satctheta] (dontknow) {Don't know};
  \node[state, right=of yes] (end1) {END};
  \node[state, right=of dontknow] (end2) {END};

  \draw (begin) edge (c);
  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
  \draw (satc) edge node[above]{NO} (yes);
  \draw (yes) edge (end1);
  \draw (satc) edge node[right]{YES} (modelc);
  \draw (modelc) edge (satctheta);
  \draw (satctheta) edge node[above]{NO} (dontknow);
  \draw (dontknow) edge (end2);
  \draw (satctheta) edge node[right]{YES} (modelctheta);
  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
\end{tikzpicture}
\end{center}

\begin{verbatim}
* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
  1. S is a trap in the net N
  2. S is marked in the initial marking M0
  3. S is unmarked in the assignment A

* Property of A_\theta: for each place s, A_\theta(s) iff s \in S

* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
  1. A ^ \delta is unsat (\delta excludes A)
  2. N |= \delta is sat  (\delta is a property of N)
\end{verbatim}

\newpage
\begin{verbatim}
* Example
  - State constraints C and \neg P are the same as in the example for method Safety.

  - Assignment A:
    x = 0
    y = 0
    z = 1
    w = 1
    t = 1

  - Trap conditions C_\theta:

    - Trap implications:
              t
         -----------
    x => y \/ z \/ w
    ^    ^    |    |
    |    |    |    w \in S
    |    |    |
    |    |    z \in S
    |    |
    |    y \in S
    |
    x \in S

    y => y \/ z \/ w
    z => true
    w => true

    - At least one of the initially marked places belongs to S:
    x

    - None of the marked places in A belongs to S:
    ~z /\ ~w
\end{verbatim}

\newpage

\begin{verbatim}
  - Assignment A_\theta:
    x = true
    y = true
    z = false
    w = false

  - Refinement constraint \delta:
    x + y >= 1
    ^   ^
    |   |
    -----
      |
      S = {x, y}, therefore \delta excludes assignment A in the next iteration

* Trap implication:

  place s \in S =>    /\       \    /
                     /  \       \  /     place p \in S
                    /    \       \/
                   t \in s*   p \in t*

  "if s is in trap S then for each output transition t at least one successor p is in trap S"

* Refinement constraint \delta:

  \Sigma s >= 1  
  A(s)

  "At least one place in S is always marked"

\end{verbatim}

\newpage

\begin{verbatim}

* Subprocedure TrapConditions:

  Input:
    (S, T, E, M0) : Petri net
    A             : Satisfying assignment for C \cup { ~P }
  Output:
    C_\theta      : Trap conditions

  Pseudocode:

\end{verbatim}
\begin{align*}
  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
    \right) \right) \land \\
  & \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land \\
  & \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right) \\
\end{align*}
\begin{verbatim}

* Subprocedure \Delta:

  Input:
    A_\theta      : Satisfying assignment for C_\theta
  Output:
    \delta        : Refinement constraint \delta

  Pseudocode:

\end{verbatim}
\begin{align*}
  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)\\
\end{align*}

\iffalse

\newpage
\section{Method Invariant}

The method Invariant constructs an invariant \verb=I= for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.

\begin{verbatim}
* Code, property, and Petri net: same as in section Method Safety.
\end{verbatim}

\begin{verbatim}
* Method Invariant

  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
  Subprocedure Model returns assignment A such that A satisfies C'.
  Subprocedure Inv constructs invariant I corresponding to N and A.
\end{verbatim}

\begin{center}
  \begin{tikzpicture}
    \node[state] (begin) {BEGIN};
    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
    \node[print, right=of satc] (noinv) {No Invariant};
    \node[action, below=of satc] (inv) {$A:=\text{Model}(C')$\\
      $I := \text{Inv}(N, A)$};
    \node[print, below=of inv] (printinv) {Invariant: $I$};
    \node[state, right=of noinv] (end1) {END};
    \node[state, below=of printinv] (end2) {END};

    \draw (begin) edge (c);
    \draw (c) edge (satc);
    \draw (satc) edge node[above]{NO} (noinv);
    \draw (satc) edge node[right]{YES} (inv);
    \draw (noinv) edge (end1);
    \draw (inv) edge (printinv);
    \draw (printinv) edge (end2);
  \end{tikzpicture}
\end{center}

\begin{verbatim}
* Property of dual state constraints C':  If C' is sat then N |= P.

* Property of invariant I:
  - I is reachable: For each reachable marking M, I(M) is valid
  - I is safe:      For markings that violate the property, I(M) is unsat
  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
\end{verbatim}

\newpage
\begin{verbatim}
* Example:

  - With weakened property P \/ (\neg \delta)

  - Negated property is \neg P /\ \delta

  - Dual state constraints C':
    0 >= (- cs1) + id1 + idle1 + (- nid1)
    0 >= (- id1) + (- idle1) + nid1 + req1
    0 >= cs1 + (- req1)
    0 >= after_you2 + (- req2)
    0 >= (- after_you2) + await2 + id2
    0 >= (- await2) + idle2
    0 >= cs2 + (- req2)
    0 >= (- cs2) + id2 + idle2
    0 >= (- id2) + (- idle2) + req2

    id1 + id2 + idle1 + idle2 < 2 * target_1 + trap_1

    id1        >= trap_1
    cs1        >= target_1
    idle1      >= 0
    req1       >= trap_1
    nid1       >= 0
    id2        >= trap_1
    cs2        >= target_1
    idle2      >= 0
    req2       >= trap_1
    after_you2 >= trap_1
    await2     >= 0
    target_1   >= 0
    trap_1     >= 0

  - Model A:

    id1        = 1
    cs1        = 1
    idle1      = 0
    req1       = 1
    nid1       = 0
    id2        = 1
    cs2        = 1
    idle2      = 0
    req2       = 1
    await2     = 0
    after_you2 = 1
    target_1   = 1
    trap_1     = 1

  - Invariant:

    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2

  - Invariant with trap constraint \delta added:

    id1 + id2 + req1 + req2 + after_you2             >= 1
    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2

\end{verbatim}

\newpage
\begin{verbatim}
* Derivation of invariant I for Petri net N that satisfies property P:

  M                                    : marking          ~ places
  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
  X                                    : firing vector    ~ transitions

  The following constraints C1 are unsat.
  M = M0 + C*X                         : place equations
  M >= 0                               : non-negativity conditions for places
  X >= 0                               : non-negativity conditions for transitions
  AM >= b                              : property P negated

  Substitute M to obtain constraints C2.
  A(M0 + CX) >= b                      : property P negated
  M0 + CX >= 0                         : non-negativity conditions for places
  X >= 0                               : non-negativity conditions for transitions

  Rewrite each system to obtain constraints C3.
  (-A*C)      (A*M0-b)                 : property P negated
  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
  (  -I)      (   0  )                 : non-negativity conditions for transitions

  Apply Farkas' Lemma to obtain constraints C4.
  yT*(-A*C)                            : 
     (  -C) = 0                        : 
     (  -I)                            :

  yT*(A*M0-b)                          : 
     (  M0  ) < 0                      : 
     (   0  )                          :

  y >= 0                               :

  The constraints C4 are sat iff the following constraints C5 are sat.
  y1 * A * C + y2 * C + y3 = 0         : 
  y1 * (A*M - b) + y2 * M0 < 0         : 
  y1 >= 0                              : 
  y2 >= 0                              : 
  y3 >= 0                              :

  The constraints C5 are sat iff the following constraints C6 are sat.
  (y1 * A + y2) * C  <= 0              : 
  (y1 * A + y2) * M0 < y1 * b          : 
  y1 >= 0                              : 
  y2 >= 0                              :

  The constraints C6 are sat iff the following constraints C' are sat.
  \lambda * C  <= 0                    : inductivity constraint
  \lambda * M0 < y1 * b                : safety constraint
  \lambda >= y1 * A                    : property constraint
  y1 >= 0                              : non-negativity constraint

  For \lambda satisfying C' the invariant is the following.
  I(M) = (\lambda * M <= \lambda * M0) : invariant
\end{verbatim}

\newpage
\begin{verbatim}
* Subprocedure \mathcal{C'}:

  Input:
    (S, T, E, M0) : Petri net
    P             : Property
  Output:
    C'            : Dual state constraints

Input:
 N = (S, T, E, M0)
   = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
 P = ( cs1 + cs2 <= 1 )

TODO: Unroll pseudocode?

* Subprocedure Inv:

  Input:
    (S, T, E, M0) : Petri net
    A             : Satisfying assignment for C'

  Output:
    I             : Invariant

TODO: Unroll pseudocode?
\end{verbatim}

\newpage
\section{Method Invariant with Minimization}

The method Invariant with Minimization constructs an invariant \verb=I=
that uses a minimal number of places for given Petri net \verb=N= and
\verb=P= when \verb=N= never violates \verb=P=. 
We present the method Invariant with Minimization by example on a
simplified mutual exclusion algorithm by Delzanno2001.

\begin{verbatim}
* Code, property, and Petri net: same as in section Method Safety.
\end{verbatim}

\begin{verbatim}
* Method Invariant with Minimization

  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A
\end{verbatim}

\begin{center}
  \begin{tikzpicture}
    \node[state] (begin) {BEGIN};
    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
    \node[print, right=of satc] (noinv) {No Invariant};
    \node[action, below=of satc] (model1) {$A:=\text{Model}(C')$};
    \node[action, below=of model1] (cm) {$C_M:=\text{MinConstraints}(N, A)$};
    \node[decision, below=of cm] (satcm) {$\text{SAT}(C' \cup C_M)$};
    \node[action, right=of satcm] (model2) {$A:=\text{Model}(C' \cup C_M)$};
    \node[action, below=of satcm] (inv) {$I := \text{Inv}(N, A)$};
    \node[print, below=of inv] (printinv) {Invariant: $I$};
    \node[state, right=of noinv] (end1) {END};
    \node[state, below=of printinv] (end2) {END};

    \draw (begin) edge (c);
    \draw (c) edge (satc);
    \draw (satc) edge node[above]{NO} (noinv);
    \draw (noinv) edge (end1);
    \draw (satc) edge node[right]{YES} (model1);
    \draw (model1) edge coordinate[pos=.5] (edgein) (cm);
    \draw (cm) edge (satcm);
    \draw (satcm) edge node[above]{YES} (model2);
    \draw (model2) |- (edgein);
    \draw (satcm) edge node[right]{NO} (inv);
    \draw (inv) edge (printinv);
    \draw (printinv) edge (end2);
  \end{tikzpicture}
\end{center}

\begin{verbatim}
* Property of minimization constraints C_M generated from A: 
    If A' satisfies C' u C_M, then Inv(N, A') uses less places than Inv(N, A)
\end{verbatim}

\newpage
\begin{verbatim}
* Example:
  - Dual state constraints C' and satisfying assignment A for C' as in example for method Invariant

  - Minimization constraints:

    nc  ~= 0 => m_nc = 1
    |           --------
    |              |
    |           place appears in invariant
    |
    place coefficient in the invariant

    nc   = 0 => m_nc  = 0
    |           ---------
    |              |
    |           place does not appear in invariant
    |
    place coefficient in the invariant

    m   ~= 0 => m_m   = 1
    m    = 0 => m_m   = 0
    cs1 ~= 0 => m_cs1 = 1
    cs1  = 0 => m_cs1 = 0
    cs2 ~= 0 => m_cs2 = 1
    cs2  = 0 => m_cs2 = 0

    m_nc + m_m + m_cs1 + m_cs2 < 3
    --------------------------   |
                |                |
                |    number of places appearing in current invariant for A
                |    = #{s | A(s) ~= 0} = #{m, cs1, cs2}
                |
                |
  number of places appearing in new invariant
\end{verbatim}

\begin{verbatim}
* Subprocedure \text{MinConstraints}:

  Input:
    (S, T, E, M0) : Petri net
    A             : Assignment
  Output:
    C_M           : Minimization Constraints

Input:
 N   = (S, T, E, M0)
     = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
 A   = ( nc=0, m=1, cs1=1, cs2=1, target1=1 )

TODO: Unroll pseudocode?

\end{verbatim}

\newpage
\section{Benchmarks of \bfc}

We experimented on the benchmarks \ttt{cprover-PN} and
\ttt{cprover\_software\_analysis} available at \\
\ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
Table~\ref{bfc-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{7cm} | r | r | r | r | } %
    \hline
    Method & pos & neg & don't know & TO \\
    \hline
    Safety (over real numbers or naturals)      & 23 &  0 & 54 &  0 \\
    Invariant \& Invariant with Minimization    & 23 &  0 & 54 &  0 \\
    Safety by Refinement                        & 23 &  0 & 54 &  0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \bfc}
\label{bfc-experiments}
\end{table}

\section{Benchmarks of \mist}

We experimented on the benchmarks available at \\
\ttt{https://github.com/pierreganty/mist/tree/master/examples/boundedPN}
and \\
\ttt{https://github.com/pierreganty/mist/tree/master/examples/PN}.
The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
Table~\ref{mist-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{6cm} | r | r | r | r | }
    \hline
    Method & pos & neg & don't know & TO \\
    \hline
    Safety (over real numbers or naturals)      & 14 &  0 & 13 &  0 \\
    Invariant \& Invariant with Minimization    & 14 &  0 & 13 &  0 \\
    Safety by Refinement                        & 20 &  0 &  7 &  0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \mist}
\label{mist-experiments}
\end{table}

\section{Benchmarks of \iic}

We experimented on the benchmarks available at \\
\ttt{https://svn.rse.mpi-sws.org/RSE/IC3forWSTS/code/trunk/examples/soter}.
The set of benchmarks consists 38 benchmarks, 28 positive and 10 negative.
Table~\ref{iic-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{6cm} | r | r | r | r | }
    \hline
    Method & pos & neg & don't know & TO \\
    \hline
    Safety (over real numbers or naturals)      & 2 & 0 & 36 & 0 \\
    Invariant \& Invariant with Minimization    & 2 & 0 & 26 & 0 \\
    Safety by Refinement                        & 2 & 0 & 36 & 0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \iic}
\label{iic-experiments}
\end{table}

We constructed the following two invariants by applying the method
Invariant with Minimization.

\begin{itemize}
\item \verb?firewall__no_pred_called_with_zero__depth_1__fixed?: \verb?(assert (<= l739 0))?
\item \verb?firewall__no_pred_called_with_zero__depth_2__fixed?: \verb?(assert (<= l3849 0))?
\end{itemize}

\iic\ constructs the following two invariants.

\begin{itemize}
\item \verb=firewall__no_pred_called_with_zero__depth_1__fixed.spec=: \verb=(l739 < 1)=
\item \verb=firewall__no_pred_called_with_zero__depth_2__fixed.spec=: \verb=(l3849 < 1)=
\end{itemize}

\fi

\end{document}