### 2014-02-03 paper/example-latex/abstract.tex

%

A number of recent approaches for proving program termination rely on
transition invariants -- a termination argument that can be
constructed incrementally using abstract interpretation.
These approaches use binary reachability analysis to check if a
candidate transition invariant holds for a given program.
For imperative programs, its efficient implementation can be obtained
by a reduction to reachability analysis, for which practical tools
are available.
In this paper, we show how a binary reachability analysis can be put
to work for proving termination of higher order functional programs.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: