### 2014-02-03 paper/example-latex/main.tex

% \documentclass[nocopyrightspace]{sigplanconf}
%\documentclass{llncs}
\documentclass{llncs}

\sloppy

\input{preamble}

\begin{document}

% \conferenceinfo{WXYZ '05}{date, City.} 
% \copyrightyear{2005} 
% \copyrightdata{[to be supplied]}

% \titlebanner{}
% \preprintfooter{}

% \pagestyle{headings}

\title{Binary Reachability Analysis of Higher Order Functional Programs}
%\title{Temporal Monitoring and Verification of Higher Order Functional Programs}
\author{Rusl\'{a}n Ledesma-Garza \and Andrey Rybalchenko}
\institute{Technische Universit\"at M\"unchen}

\maketitle

\begin{abstract}
%

\input{abstract}
\iffalse
The automata-theoretic approach to program verification is an
effective and uniform way of dealing with complex temporal properties.
The crux of the approach is to represent the property by a monitor
that inspects program computations.
The monitor keeps record of the (intermediate) inspection results, and
if the property violation is detected then an appropriate note is
recorded.
Then, the monitor is composed with the program, which produces a
product program that monitors its own computation and can be analyzed
using tools for proving reachability and (fair) termination (i.e.,
sophisticated temporal reasoning is no longer necessary).
While already ubiquitous for dealing with imperative programs, monitor
based approaches were recently applied for termporal verification of
higher order functional programs by using higher order recursion
schemes (HORS), e.g.~\cite{Ong06,Kobayashi09}.

In this paper, we present an alternative approach to monitoring and
verifying higher order functional programs that applies the monitor
composition directly on the program source code (instead of composing
with a HORS derived from the program by applying finite state
abstraction on data types).
We formulate our approach for a \miniocaml language as a program
transformation algorithm and present its implementation.
Our experimental evaluation in combination with the DSolve verifier
for OCaml~\cite{DSolveCAV} shows that the scalability and
effectiveness of existing abstraction-based assertion checking tools
can be effectively leveraged for dealing with temporal safety and
liveness properties of higher order functional programs.

The automata-theoretic approach is a standard method for dealing with
complex temporal properties of imperative programs and only requires
program verifiers that deal reachability and (fair) termination
properties.
The automata-theoretic semantics of imperative programs directly
supports monitoring and product construction.
Until now, the temporal reasoning about functional programs does not
rely on monitoring and reduction to reachability and (fair)
termination.
Instead, the reasoning is executed using type systems and type
inference algorithms that are developed/customized for the specific
property at hand.
\fi

\iffalse

In this paper, we present an application of the automata-theoretic
approach to the temporal verification, i.e., safety and liveness, of
\miniocaml\ programs with higher-order procedures.  
First, we present the notion of monitoring for \miniocaml\ computations
given as evaluation trees, and define monitor specifications that can be
composed with \miniocaml\ programs.
Second, we present a product construction algorithm that composes an
\miniocaml\ program with a monitor specification in the presence of
higher order procedures.
Third, we present an implementation of the proposed approach that uses
the \dsolve\ reachability checker, together with its experimental
evaluation on examples from the literature.
The examples include resource analysis and termination checking tasks.
\fi
\end{abstract}

% \mbox{ }
% \\ \\
%\nocite{Hofmann00,Hofmann02,Hofmann03,Kobayashi09}

\input{sec_introduction}
\input{sec_illustration}
\input{sec_preliminaries}
\input{sec_semantics}
%\input{sec_logger}
\input{sec_transformation}
%\input{sec_trans_assert}
\input{sec_experiments}
\input{sec_relwork.tex}
\input{sec_conclusions}

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: