### 2014-02-03 paper/example-latex/sec_experiments.tex

\section{Experimental evaluation}
\label{sec-experiments}

In this section we describe our implementation and the corresponding
experimental evaluation.

\paragraph{Implementation.}
We implemented \product\ as an extension to the \camlp\
parser~\cite{Camlp4-dist}. Our implementation takes as input a user
program and a specification consisting of a function name, an arity,
and candidate transition invariant. Our implementation produces a
transformed program following the procedure depicted in
Figure~\ref{fig-ex-prod}.

\paragraph{Experiments.}
\input{fig_experiments}
Our experiments consisted of two steps. First we applied our
transformation to the set of benchmarks summarized in
Figure~\ref{fig-experiments}. Then we analyzed the transformed
benchmarks using the reachability checker \dsolve~\cite{Dsolve, DsolveCAV}.
\ifthenelse{\equal{\isTechReport}{true}}{
The set of benchmarks is available at~\ttt{http://www7.in.tum.de/\~{}ruslan/binreach/}.
}{
The set of benchmarks is available at~\cite{FunVTechReport}.
}
Our benchmarks feature higher order functions and algebraic
data types (lists).  We summarize our verified benchmarks in
Figure~\ref{fig-experiments}.
Benchmarks 11-15 correspond to the higher-order programs
in~\cite{Sereni05terminationanalysis} that are strict and
\mbox{type-check}.
The experiments show that our transformation can be used together with a state of
the art static analyzer to prove termination of higher-order programs found
in the literature.

% \paragraph{Evaluation.}
% The experiments show that our transformation can be used together with a state of
% the art static analyzer to prove termination of higher-order programs found
% in the literature.

% For example, consider the higher-order benchmark \ttt{sereni29} shown below.
% \begin{verbatim}
% let rec length l =
%   match l with
%     | [] -> 0
%     | _ :: t -> 1 + length t
% let rec map f xs =
%   match xs with
%     | [] -> []
%     | y :: ys -> f y :: map f ys in
% let compose f g x = f (g x) in
% let add x y = x + y in
% map (compose (add 1) (add 2)) [0;1]
% \end{verbatim}
% We applied our program transformation on \ttt{sereni29} using the function
% identifier \ttt{map}, the number of parameters $2$, and the transition
% invariant candidate $\ttt{length}(a_2) \geq 0 \wedge
% \ttt{length}(m_2) > \ttt{length}(a_2)$. Then we obtained a
% transformed program that we verified correct using \dsolve.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: