### 2014-02-05 paper/workspace/sec_experiments.tex
Included experiments by Filip.

\section{Experiments}

To evaluate the described methods, we implemented them in a tool called
\emph{Petrinizer}. Petrinizer is essentially a set of Bash and Prolog scripts
that mediate between the input, specified in the input format of
MIST2\footnote{\url{https://github.com/pierreganty/mist}}, and the SMT
solver Z3\footnote{\url{http://z3.codeplex.com/}} \cite{DeMouraTACAS08}. With
the choice of Bash and Prolog over languages with bindings for Z3 API, we
restricted Z3's full power. Indeed, by using Z3 as an external tool, we
were not able to fully exploit the incremental nature of its linear arithmetic
solver (TODO: cite appropriate papers). However, even the non-optimal
implementation turned out to be robust and scalable, with the main advantage of
being able to tackle problem instances that are out of reach for
state-of-the-art coverability checkers.

The evaluation of Petrinizer had three main goals. First, we wanted to compare
its performance against state-of-the-art tools like MIST2, BFC
and IIC (cite!). Second, as the method it
implements is incomplete, we wanted to measure its rate of success on safe problem instances. As a
subgoal, we wanted to investigate the usefulness and necessity of traps, as
well as to what extent real arithmetic suffices over integer arithmetic in
proving safety. And last, since the language of linear arithmetic, used by
Petrinizer, allows for a more succint representation of formulas than the
language used by IIC, we wanted to compare sizes of invariants the two tools produce.

Inputs that were used in tests come from various sources (TODO: write a
sentence to justify the variety).
One source is the collection of Petri nets from the literature that is part of the MIST2 toolkit.
The second source are Petri nets originating from the analysis of concurrent C
programs that were used in the evaluation of BFC \cite{KaiserCONCUR12}. Then
there are inputs originating from the provenance analysis of a medical and a
bug-tracking system \cite{MajumdarSAS13}. Finally, we generated a set of Petri
nets from Erlang programs, using an Erlang verification tool Soter
\cite{DOsualdoSAS13}. (TODO: Acknowledge Emanuele D'Osualdo for help.)

All experiments were performed on the identical machines, equipped with
Intel Xeon 2.66 GHz CPUs and 48 GB of memory, running Linux 3.2.48.1 in 64-bit
mode. Execution time was limited to 100,000 seconds (27 hours, 46 minutes and
40 seconds), and memory to 2 GB.

\subsection{Performance}

Explain what we observe:
\begin{itemize}
  \item On small examples IIC mostly outperforms Petrinizer. Probably due
    to a slower parser, setting up and invoking Z3.
  \item On examples of a similar size, coming from the same source,
    unlike with IIC, resource consumption is fairly uniform. (Medical examples.)
  \item Petrinizer is able to handle some huge examples that are out of
    scope for IIC.
  \item There were no cases when IIC was able to finish, but Petrinizer ran
    out of resources. As a matter of fact, Petrinizer finishes in all cases,
    whereas other tools somethimes fail.
\end{itemize}

\emph{Invariant sizes.} Again, explain how we measure it. For IIC, we measure
the number of non-null literals, the number of clauses, total length of
the invariant in characters. For Petrinizer, we measure the number of
non-null coefficients in the linear expressions and the total length of the
invariant in characters. Argue why it makes sense to compare any of these
quantities. Point out the difference in size with and without the invariant
minimization.

\emph{Rate of success.} Point out that Petrinizer mostly successfully proves
safety. Specifically, it proves safety in all of the safe Soter examples.
Point out the cases where traps were actually useful, and potentially progress
into some pseudo-philosophical discussion on what constitutes a good set of examples.