### 2014-02-05 paper/workspace/sec_preliminaries.tex

\section{Preliminaries}

\begin{verbatim}
* Petri net:
 N = (S, T, E, M0), where
     S : places
     T : transitions
     E : edges
     M0: initial marking

* Property:
 P, linear arithmetic formula
\end{verbatim}

\iffalse
\begin{verbatim}
 Examples:
  x + y < 0
  x + y < 0 \\and x + z > 0
  x + y < 0 \\or  x + z > 0

 ~P = (x + y <  0 \\and x + z >  0) \\or  x + z >  0
  P = (x + y >= 0 \\or  x + z <= 0) \\and x + z <= 0
\end{verbatim}
\fi