### 2014-02-05 paper/workspace/fig_dia_method_safety_by_refinement.tex

\begin{figure}
  \begin{center}
    \begin{tikzpicture}[
      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
      ]
      \node[state] (begin) {BEGIN};
      \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
      \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
      \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
        $C_{\theta}:=\text{TrapConditions}(N, A)$};
      \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
      \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
        $\delta:=\Delta(A_\theta)$\\
        $C:=C \cup \{\delta\}$};
      \node[print, right=of satc] (yes) {N satisfies P};
      \node[print, right=of satctheta] (dontknow) {Don't know};
      \node[state, right=of yes] (end1) {END};
      \node[state, right=of dontknow] (end2) {END};

      \draw (begin) edge (c);
      \draw (c) edge coordinate[pos=.5] (edgein) (satc);
      \draw (satc) edge node[above]{NO} (yes);
      \draw (yes) edge (end1);
      \draw (satc) edge node[right]{YES} (modelc);
      \draw (modelc) edge (satctheta);
      \draw (satctheta) edge node[above]{NO} (dontknow);
      \draw (dontknow) edge (end2);
      \draw (satctheta) edge node[right]{YES} (modelctheta);
      \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
      -| ([xshift=-2cm] modelctheta.west) |- (edgein);
    \end{tikzpicture}
  \end{center}
  \caption{Diagram for Method Safety by Refinement}
  \label{fig:method-safety-by-refinement-diagram}
\end{figure}