### 2014-02-05 paper/workspace/fig_dia_method_invariant.tex

\begin{figure}
  \begin{center}
    \begin{tikzpicture}[
      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
      ]
      \node[state] (begin) {BEGIN};
      \node[action, below=of begin] (c) {$C':=\mathcal C'(N, \neg P)$};
      \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
      \node[print, right=of satc] (noinv) {No Invariant};
      \node[action, below=of satc] (inv) {$A':=\text{Model}(C')$\\
        $I := \text{Inv}(N, A')$};
      \node[print, below=of inv] (printinv) {Invariant: $I$};
      \node[state, right=of noinv] (end1) {END};
      \node[state, below=of printinv] (end2) {END};

      \draw (begin) edge (c);
      \draw (c) edge (satc);
      \draw (satc) edge node[above]{NO} (noinv);
      \draw (satc) edge node[right]{YES} (inv);
      \draw (noinv) edge (end1);
      \draw (inv) edge (printinv);
      \draw (printinv) edge (end2);
    \end{tikzpicture}
  \end{center}
  \caption{Diagram for Method Invariant}
  \label{fig:method-invariant-diagram}
\end{figure}