### 2014-02-06 paper/example-latex-threads/abstract.tex
Example?

%
Automated verification of multi-threaded programs requires explicit
identification of the interplay between interacting threads, so-called
environment transitions, to enable scalable, compositional reasoning.  
Once the environment transitions are identified, we can prove program
properties by considering each program thread in isolation, as the
environment transitions keep track of the interleaving with other
threads.
Finding adequate environment transitions that are sufficiently precise
to yield conclusive results and yet do not overwhelm the verifier with
unnecessary details about the interleaving with other threads is a
major challenge.
In this paper we propose a method for safety verification of
multi-threaded programs that applies (transition) predicate
abstraction-based discovery of environment transitions, exposing a
minimal amount of information about the thread interleaving.
The crux of our method is an abstraction refinement procedure that
uses recursion-free Horn clauses to declaratively state abstraction
refinement queries.
Then, the queries are resolved by a corresponding constraint solving
algorithm.
We present preliminary experimental results for mutual exclusion
protocols and multi-threaded device drivers.

\iffalse % popl submission abstract
%
In order to scale, verification of multi-threaded programs requires
definition of auxiliary variables that reveal some details about the
local state of the concurrent threads and construction of
rely/guarantee assertions that describe the thread interaction in
terms of shared and auxiliary variables.
%
In this paper we present an automatic method for the discovery of
auxiliary variables and relies/guarantees.
Our method is based on a generalization of counterexample guided
abstraction refinement to the multi-threaded setting.
The crux of the generalization is a notion of tree-like
counterexamples that explicitly account for thread interleaving.
By solving the corresponding tree-like path constraints, our method
reveals additional local facts through auxiliary variables and
iteratively strengthens relies/guarantees.
Furthermore, the tree-like path constraints can avoid the introduction
of auxiliary variables when they are not necessary.
%Furthermore, the tree-like path constraints enable the discovery of
%modular relies/guarantees (whose auxiliary variables do not reveal any
%local facts) whenever a modular proof exists.
We devised a solver for the tree-like constraints that is based on
Craig interpolation and used it in a prototype implementation of our
method.
We present preliminary experimental results for mutual exclusion
protocols and concurrent device drivers.
\fi

\iffalse % PLDI abstract
Automated verification of multi-threaded programs requires explicit
identification of the interplay between interacting threads, so-called
environment assumptions, to enable scalable reasoning.  
Once identified, these assumptions can be used for reasoning with one
program thread at a time, which is possible by using the respective
environment assumption to model the interleaving with other threads.
Finding adequate assumptions that are sufficiently precise to yield
conclusive results and yet keep track only of necessary facts about
the execution environment in order to scale well is a major challenge.
In this paper we propose a constraint-based technique for the
inference of such assumptions.
Our technique automatically steers towards an optimal
precision/efficiency trade-off between the extremes of efficient, but
incomplete thread-modular reasoning and complete, but prohibitively
expensive consideration of all interleavings.
For this task, we pinpoint a declarative formulation of modular
verification that allows one to express requisite environment
assumptions using constraints and admits algorithmic solutions based
on abstract fixpoint checking.
We describe an application of our environment assumption inference for
the verification of reachability and termination properties of
multi-threaded programs, and present our experience with its
implementation as well as evaluation in practice. 
\fi

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: