### 2014-02-06 paper/example-latex-threads/popl213af-gupta.tex

\documentclass{sigplanconf}
\sloppy

\usepackage{graphicx,pgf,tikz,tikz-qtree}
\usepackage{macros,ar-alg}
\usepackage{amsmath,amssymb,xspace}
%\usepackage{mathabx} % provides \smalltriangleright
\usepackage{color} % graphicx scaprs ams symbols and we don't use it anyway
%\usepackage{extarrows}
\usetikzlibrary{arrows,automata}
\usepackage{ifthen}

\newcommand{\isTechReport}{true} % true or false
\newcommand{\includeProof}[1]{
  \ifthenelse{\equal{\isTechReport}{true}}{
    #1
  }{
  }
}

\begin{document}

\conferenceinfo{POPL'11,} {January 26--28, 2011, Austin, Texas, USA.}
\CopyrightYear{2011}
\copyrightdata{978-1-4503-0490-0/11/01}

\titlebanner{}        % These are ignored unless
\preprintfooter{}   % 'preprint' option specified.

%\title{Predicate Abstraction and Refinement with Horn Clauses for Verifying Multi-Threaded Programs}
\title{Predicate Abstraction and Refinement for Verifying Multi-Threaded Programs}
%\title{Constraint Based Iterative Inference of Auxiliary Variables 
%  and Relies/Guarantees}
%\title{Constrained Environment Inference for\\ 
% Verification of Multi-Threaded Programs}
%\title{Constraint Based Inference of Auxiliary Variables 
%  and Relies/Guarantees for Verifying Multi-Threaded Programs}
% \subtitle{Draft. Please don't distribute.}

\authorinfo{Ashutosh Gupta \and Corneliu Popeea \and Andrey Rybalchenko}
           {Institut f\"ur Informatik, Technische Universit\"at M\"unchen\\Germany}
           {\{guptaa,popeea,rybal\}@in.tum.de}
%\authorinfo{Corneliu Popeea}{\TUM}{}
%\authorinfo{Andrey Rybalchenko}{\TUM}{}

\maketitle

\category{D.2.4}{Software Engineering}{Software/Program Verification}
\category{F.3.1}{Logics and Meanings of Programs}{Specifying and Verifying and Reasoning about Programs}
\terms
Algorithms, Languages, Reliability, Verification.
\keywords
Multi-threaded programs, safety, proof rule,
modular reasoning, environment transitions, (transition) predicate
abstraction, abstraction refinement, Horn clauses.

% TBD in final version
\vspace{-0.3cm}
\acks

The first author was supported by the DFG Graduiertenkolleg 1480 (PUMA).
We thank Byron Cook, Rusl\'an Ledesma Garza, and Peter O'Hearn for
comments and suggestions.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.
%%%%%%% \bibliography{biblio}

\begin{thebibliography}{34}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Bar-David and Taubenfeld(2003)]{BarDavidDISC03}
Y.~Bar-David and G.~Taubenfeld.
\newblock Automatic discovery of mutual exclusion algorithms.
\newblock In \emph{DISC}, pages 136--150, 2003.

\bibitem[Basler et~al.(2009)Basler, Mazzucchi, Wahl, and Kroening]{BaslerCAV09}
G.~Basler, M.~Mazzucchi, T.~Wahl, and D.~Kroening.
\newblock Symbolic counter abstraction for concurrent software.
\newblock In \emph{CAV}, pages 64--78, 2009.

\bibitem[Basler et~al.(2010)Basler, Hague, Kroening, Ong, Wahl, and
  Zhao]{BaslerTACAS10}
G.~Basler, M.~Hague, D.~Kroening, C.-H.~L. Ong, T.~Wahl, and H.~Zhao.
\newblock Boom: Taking boolean program model checking one step further.
\newblock In \emph{TACAS}, pages 145--149, 2010.

\bibitem[Clarke et~al.(2000)Clarke, Grumberg, Jha, Lu, and Veith]{ClarkeCEGAR}
E.~M. Clarke, O.~Grumberg, S.~Jha, Y.~Lu, and H.~Veith.
\newblock Counterexample-guided abstraction refinement.
\newblock In \emph{CAV}, pages 154--169, 2000.

\bibitem[Cohen and Namjoshi(2009)]{CohenFMSD09}
A.~Cohen and K.~S. Namjoshi.
\newblock Local proofs for global safety properties.
\newblock \emph{FMSD}, 34\penalty0 (2):\penalty0 104--125, 2009.

\bibitem[Corbet et~al.(2005)Corbet, Rubini, and Kroah-Hartman]{CorbetBook05}
J.~Corbet, A.~Rubini, and G.~Kroah-Hartman.
\newblock \emph{Linux Device Drivers, 3rd Edition}.
\newblock O'Reilly Media, Inc., 2005.

\bibitem[Cousot and Cousot(1977)]{lattice77}
P.~Cousot and R.~Cousot.
\newblock Abstract interpretation: A unified lattice model for static analysis
  of programs by construction or approximation of fixpoints.
\newblock In \emph{POPL}, pages 238--252, 1977.

\bibitem[Flanagan and Godefroid(2005)]{FlanaganPOPL05}
C.~Flanagan and P.~Godefroid.
\newblock Dynamic partial-order reduction for model checking software.
\newblock In \emph{POPL}, pages 110--121, 2005.

\bibitem[Flanagan and Qadeer(2003)]{ThreadModularFlanaganQadeer}
C.~Flanagan and S.~Qadeer.
\newblock Thread-modular model checking.
\newblock In \emph{SPIN}, pages 213--224, 2003.

\bibitem[Flanagan et~al.(2002)Flanagan, Freund, and Qadeer]{CalvinESOP}
C.~Flanagan, S.~N. Freund, and S.~Qadeer.
\newblock Thread-modular verification for shared-memory programs.
\newblock In \emph{ESOP}, pages 262--277, 2002.

\bibitem[Godefroid(1994)]{GodefroidPhd94}
P.~Godefroid.
\newblock \emph{Partial-Order Methods for the Verification of Concurrent
  Systems - An Approach to the State-Explosion Problem}.
\newblock PhD thesis, University of Liege, Computer Science Department, 1994.

\bibitem[Graf and Sa\"{\i}di(1997)]{GrafSaidi}
S.~Graf and H.~Sa\"{\i}di.
\newblock Construction of abstract state graphs with {PVS}.
\newblock In \emph{CAV}, pages 72--83, 1997.

\bibitem[Gupta et~al.(2010)Gupta, Popeea, and Rybalchenko]{GuptaATVA10}
A.~Gupta, C.~Popeea, and A.~Rybalchenko.
\newblock Non-monotonic refinement of control abstraction for concurrent
  programs.
\newblock In \emph{ATVA}, pages 188--202, 2010.

\bibitem[Henzinger et~al.(2002)Henzinger, Jhala, Majumdar, and
  Sutre]{BlastLazy02}
T.~A. Henzinger, R.~Jhala, R.~Majumdar, and G.~Sutre.
\newblock Lazy abstraction.
\newblock In \emph{POPL}, pages 58--70, 2002.

\bibitem[Henzinger et~al.(2004)Henzinger, Jhala, and Majumdar]{HenzingerPLDI04}
T.~A. Henzinger, R.~Jhala, and R.~Majumdar.
\newblock Race checking by context inference.
\newblock In \emph{PLDI}, pages 1--13, 2004.

\bibitem[Jones(1983{\natexlab{a}})]{JonesTOPLAS83}
C.~B. Jones.
\newblock Tentative steps toward a development method for interfering programs.
\newblock \emph{ACM Trans. Program. Lang. Syst.}, 5\penalty0 (4):\penalty0
  596--619, 1983{\natexlab{a}}.

\bibitem[Jones(1983{\natexlab{b}})]{RelyGuarantee}
C.~B. Jones.
\newblock Specification and design of (parallel) programs.
\newblock In \emph{IFIP Congress}, pages 321--332, 1983{\natexlab{b}}.

\bibitem[Lamport(1974)]{LamportCACM74}
L.~Lamport.
\newblock A new solution of {D}ijkstra's concurrent programming problem.
\newblock \emph{Commun. ACM}, 17\penalty0 (8):\penalty0 453--455, 1974.

\vfill\eject

\bibitem[Lamport(1987)]{LamportTCS87}
L.~Lamport.
\newblock A fast mutual exclusion algorithm.
\newblock \emph{ACM Trans. Comput. Syst.}, 5\penalty0 (1):\penalty0 1--11,
  1987.

\bibitem[Lu et~al.(2008)Lu, Park, Seo, and Zhou]{LuASPLOS08}
S.~Lu, S.~Park, E.~Seo, and Y.~Zhou.
\newblock Learning from mistakes: a comprehensive study on real world
  concurrency bug characteristics.
\newblock In \emph{ASPLOS}, pages 329--339, 2008.

\bibitem[Malkis et~al.(2006)Malkis, Podelski, and Rybalchenko]{MalkisICTAC06}
A.~Malkis, A.~Podelski, and A.~Rybalchenko.
\newblock Thread-modular verification is cartesian abstract interpretation.
\newblock In \emph{ICTAC}, pages 183--197, 2006.

\bibitem[Manna and Pnueli(1995)]{MannSafetyBook95}
Z.~Manna and A.~Pnueli.
\newblock \emph{Temporal verification of reactive systems: safety}.
\newblock Springer-Verlag, 1995.

\bibitem[McKenney(2007)]{McKenneyLWN07}
P.~McKenney.
\newblock Using {Promela} and {Spin} to verify parallel algorithms.
\newblock \emph{LWN.net weekly edition}, 2007.

\bibitem[Musuvathi et~al.(2008)Musuvathi, Qadeer, Ball, Basler, Nainar, and
  Neamtiu]{MusuvathiOSDI08}
M.~Musuvathi, S.~Qadeer, T.~Ball, G.~Basler, P.~A. Nainar, and I.~Neamtiu.
\newblock Finding and reproducing {Heisenbugs} in concurrent programs.
\newblock In \emph{OSDI}, pages 267--280, 2008.

\bibitem[Necula et~al.(2002)Necula, McPeak, Rahul, and Weimer]{NeculaCC02}
G.~C. Necula, S.~McPeak, S.~P. Rahul, and W.~Weimer.
\newblock {CIL}: Intermediate language and tools for analysis and
  transformation of {C} programs.
\newblock In \emph{CC}, pages 213--228, 2002.

\bibitem[Owicki and Gries(1976)]{OwickiAI76}
S.~S. Owicki and D.~Gries.
\newblock An axiomatic proof technique for parallel programs {I}.
\newblock \emph{Acta Inf.}, 6:\penalty0 319--340, 1976.

\bibitem[Pnueli et~al.(2002)Pnueli, Xu, and Zuck]{PnueliCAV02}
A.~Pnueli, J.~Xu, and L.~D. Zuck.
\newblock Liveness with (0, 1, infty)-counter abstraction.
\newblock In \emph{CAV}, pages 107--122, 2002.

\bibitem[Podelski and Rybalchenko(2005)]{TransPred05}
A.~Podelski and A.~Rybalchenko.
\newblock Transition predicate abstraction and fair termination.
\newblock In \emph{POPL}, pages 132--144, 2005.

\bibitem[Podelski and Rybalchenko(2007)]{RybalPodelskiPADL07}
A.~Podelski and A.~Rybalchenko.
\newblock {ARMC}: The logical choice for software model checking with
  abstraction refinement.
\newblock In \emph{PADL}, pages 245--259, 2007.

\bibitem[Qadeer and Wu(2004)]{QadeerPLDI04}
S.~Qadeer and D.~Wu.
\newblock {KISS}: keep it simple and sequential.
\newblock In \emph{PLDI}, pages 14--24, 2004.

\bibitem[Rybalchenko()]{RybalchenkoARMC}
A.~Rybalchenko.
\newblock {The ARMC tool}.
\newblock \emph{Available from http://www7.in.tum.de/\~{}rybal/armc/}.

\bibitem[Szymanski(1988)]{SzymanskiICS88}
B.~K. Szymanski.
\newblock A simple solution to {L}amport's concurrent programming problem with
  linear wait.
\newblock In \emph{ICS}, pages 621--626, 1988.

\bibitem[{The Intelligent Systems Laboratory}(2001)]{sicstus}
{The Intelligent Systems Laboratory}.
\newblock \emph{SICStus Prolog User's Manual}.
\newblock Swedish Institute of Computer Science, 2001.
\newblock Release 3.8.7.

\bibitem[Wang et~al.(2008)Wang, Yang, Kahlon, and Gupta]{WangTACAS08}
C.~Wang, Z.~Yang, V.~Kahlon, and A.~Gupta.
\newblock Peephole partial order reduction.
\newblock In \emph{TACAS}, pages 382--396, 2008.

\end{thebibliography}

\iffalse
\appendix
% \input{auxiliaries}
% \input{redundancy}
% \input{algorithms}
% \input{tables}

\clearpage
\input{fig-alg-main}
\input{fig-alg-abst-reach}
\input{fig-alg-cex-path}
\input{fig-alg-refine}
\input{fig-alg-mkHornClauses}
\input{fig-alg-solveHorn}
\fi
%\clearpage
%\input{fig-ex-mk-tree}

\end{document}