### 2014-03-17 paper/workspace/appendix.tex
Added submitted version of paper

\section{Proof of Theorem \ref{thm:dual}}

We use the following version of Farkas' Lemma:

\begin{theorem}[Farkas' lemma (variant) \cite{Schrijver86}]
\label{lem:farkas}
Let $A$ be a matrix and let $b$ be a vector.
Then the system $Ax \le b$ of linear inequalities has a solution $x$,
if and only if $yb \ge 0$ for each row vector $y \ge 0$ with $yA = 0$.
\end{theorem}

\begin{reftheorem}{thm:dual}
The system $\mathcal{S}'$ of constraints given by
\begin{align*}
  \lambda C    & \le 0            && \text{inductivity constraint} \\
  \lambda m_0  & <   Y_1 b + Y_2 1 && \text{safety constraint} \\
  \lambda      & \ge Y_1 A + Y_2 D && \text{property constraint} \\
  Y_1, Y_2     & \ge 0     && \text{non-negativity constraint} 
\end{align*}
is feasible over the rational numbers if{}f the system $\mathcal{S}$ is infeasible.
\end{reftheorem}
\begin{proof}
Recall that the system $\mathcal{S}$ is given by:
\begin{align*}
  \mathcal{C}&(P, T, F, m_0)  && \text{state constraints} \\
  A M & \ge b                 && \text{negation of property $P$} \\
  D M & \ge 1                 && \text{trap constraints}
\end{align*}
Eliminating $M$ we get
\begin{align*}
A (m_0 + C X) & \ge b && \text{property $P$ negated} \\
D (m_0 + C X) & \ge 1 && \text{trap constraints} \\
   m_0 + C X  & \ge 0 && \text{non-negativity conditions for places} \\
           X  & \ge 0 && \text{non-negativity conditions for transitions}
\end{align*}
which can be rewritten in matrix form as
%
\begin{align*}
\begin{pmatrix}
  -  A  C \\
  -  D  C \\
  -     C \\
  -     I
\end{pmatrix} X & \le
\begin{pmatrix}
  A m_0 - b \\
  D m_0 - 1 \\
    m_0     \\
     0
\end{pmatrix}
\end{align*}
%
Lemma \ref{lem:farkas} yields the set $\mathcal{S}_1$ given by  
%
\begin{align*}
  Y^T
\begin{pmatrix}
  -  A  C \\
  -  D  C \\
  -     C \\
  -     I
\end{pmatrix} & = 0  &
  Y^T
\begin{pmatrix}
  A m_0 - b \\
  D m_0 - 1 \\
    m_0     \\
     0
\end{pmatrix} & < 0  &
Y & \ge 0
\end{align*}
So, by the Lemma, $\mathcal{S}_1$ is satisfiable if and only if $\mathcal{S}$ is unsatisfiable. 
Splitting the vector $Y$ into four segments $Y := (Y_1|Y_2|Y_3|Y_4)$ 
we obtain that $\mathcal{S}_1$ is equisatisfiable with 
\begin{align*}
  Y_1 A C         + Y_2 D C         + Y_3 C  + Y_4    & = 0 \\
  Y_1 (A m_0 - b) + Y_2 (D m_0 - 1) + Y_3 m_0         & < 0 \\
  Y_i                               & \ge 0 \ \ i=1,2,3,4
\end{align*}
and, since $Y_4\ge 0$ is only requiring that $Y_1 D C + Y_2 A C + Y_3 C \leq 0$ holds,
also with 
\begin{align*}
  (Y_1 A + Y_2 D + Y_3) C    & \le 0          \\
  (Y_1 A + Y_2 D + Y_3) m_0  & < Y_1 b + Y_2 1 \\
  Y_1,Y_2,Y_3 & \ge 0 
\end{align*}
Finally, defining $\lambda := Y_1 A + Y_2 D + Y_3$, we obtain that
$\mathcal{S}_1$ is equisatisfiable with $\mathcal{S}'$:
\begin{align*}
  \lambda C    & \le 0             \\
  \lambda m_0  & <   Y_1 b + Y_2 1  \\
  \lambda      & \ge Y_1 A + Y_2 D  \\
  Y_1, Y_2        & \ge 0    
\end{align*}
\end{proof}