### 2014-03-17 paper/workspace/fig_safety_by_refinement.tex

\begin{figure}[t]
  \begin{algorithmic}[1]
    \State $C := \mathcal C(N)$
    \While {SAT$(C \cup \{\neg P\})$}
    \State $A := $Model$(C \cup \{\neg P\})$
    \State $C_\theta := $TrapConditions$(N, A)$
    \If {SAT$(C_\theta)$}
    \State $A_\theta := $Model$(C_\theta)$
    \State $\delta := \Delta(A_\theta)$
    \State $C := C \cup \{\delta\}$
    \Else
    \State \Return 'The petri net may not satisfy the property'
    \EndIf
    \EndWhile
    \State \Return 'The petri net satisfies the property!'
  \end{algorithmic}
  \caption{Pseudocode for Function Safety by Refinement}
  \label{fig_function_safety_by_refinement_pseudocode}
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: