### 2014-03-17 paper/workspace/main.tex

\documentclass{llncs}

\newcommand{\isDraft}{false}

\input{preamble}

\ifthenelse{\equal{\isDraft}{true}}{
  \usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
  \pagestyle{plain}
}{}

\begin{document}

\nocite{DOsualdoSAS13}

\ifthenelse{\equal{\isDraft}{false}}{
  \title{An SMT-based Approach to\\
         Coverability Analysis}
  \author{
    Javier~Esparza\inst{1}
    \and
    Rusl\'{a}n~Ledesma-Garza\inst{1}
    \and
    Rupak~Majumdar\inst{2}
    \and
    Philipp~Meyer\inst{1}
    \and
    Filip~Niksic\inst{2}
  }
  \institute{
    Technische Universit\"at M\"unchen
    \and
    Max Planck Institute for Software Systems (MPI-SWS)
  }

  \maketitle

  \begin{abstract}
    Model checkers based on Petri net coverability have been used
    successfully in recent years to verify safety properties of concurrent
    shared-memory or asynchronous message-passing software. We revisit 
    a constraint approach to the problem based on
    classical Petri net analysis techniques. We show how to 
    utilize an SMT solver to implement
    the constraint approach, and additionally, to generate a minimized
    inductive invariant from a safety proof. We empirically evaluate 
    our procedure on a large set of existing Petri
    net benchmarks. Even though our technique is incomplete, 
    it can quickly discharge all safe instances. Additionally, the
    inductive invariants computed are usually orders of magnitude smaller
    than those produced by existing solvers.
  \end{abstract}
}{
}

\input{sec_introduction}
\input{sec_rel_work}
\input{sec_preliminaries}
\input{sec_safety}
\input{sec_safety_by_refinement}
%\input{sec_invariant}
\input{sec_invariant_by_refinement}
\input{sec_invariant_w_minimization}
\input{sec_experiments}
% \input{sec_conclusions}

\bibliographystyle{abbrv}
\bibliography{references}

\newpage
\appendix
\input{appendix}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: