### 2014-03-17 paper/workspace/sec_introduction.tex

\section{Introduction}

In recent years many papers have proposed and developed techniques for
the verification of concurrent software
\cite{KaiserKW10,GantyM12,BouajjaniEmmi12,KaiserKWCONCUR12,DOsualdoSAS13}.
In particular, model checkers based on Petri net coverability have been successfully
applied.  Petri nets are a natural model for this problem for two
resons.  The first and obvious one is that they are a simple and
natural automata-like model for concurrent systems. The second, more
specific, is that they can model certain programs with an unbounded
number of threads or thread creation. In a nutshell, the places of the
net correspond to program points, and the number of tokens in a place
models the number of threads that are currently at that point. This
point, which was first observed in \cite{GS92}, was later revisited in
\cite{DelzannoRB02}, and, more implicitly, in \cite{KaiserKW10}.

The problem whether at least one thread can reach a given program
point (modelling some kind of error), naturally reduces to the {\em
  coverability problem} for Petri nets: given a net $N$ and a marking
$M$, decide whether some reachable marking of $N$ {\em covers} $M$,
i.e., puts at least as many tokens as $M$ on each place. While the
decidability and EXPSPACE-completeness of the coverability problem
were settled long ago \cite{KarpM69,Rackoff78}, new algorithmic ideas
have been developed in recent years
\cite{GeeraertsRB06,GantyFI08,ValmariH12,KaiserKWCONCUR12,KloosMNP13}. 
The techniques are based on forward or backward state-space exploration, 
which is accelerated in a number of ways in order to cope with the possibly 
infinite number of states.

In this paper we revisit an approach to the coverability problem based
on classical Petri net analysis techniques: the marking equation and
traps \cite{murata1989petri,EsparzaM00,Reisig13}.
The marking equation is a system of linear constraints that can be
easily derived from the net, and whose set of solutions
overapproximates the set of reachable markings.  The marking equation
can be supplemented with linear constraints specifying a set of unsafe
markings, and solved using standard linear or integer programming. If
the equations are infeasible, then all reachable markings are safe. If
not, then, as suggested in \cite{EsparzaM00}, one can use a
counterexample to derive an additional constraint in the shape of a
{\em trap}: a set of places that, loosely speaking, once marked cannot
be ``emptied''.

While these techniques are well known, their application has always
suffered from the absence of efficient decision procedures for 
linear arithmetic together with Boolean satisfiability.  
Moreover, and to the best of our knowledge, their
systematic application to verification of concurrent software has not
been studied yet. Profiting from recent advances in SMT-solving
technology, we reimplement the techniques on top of the Z3 SMT solver
\cite{MouraB08}, and apply them to a large collection of benchmarks.

Our first and surprising finding is that, while the technique is theoretically not
complete, it is powerful enough to prove safety of 96 out of a total
of 115 safe benchmarks gathered from current research papers in concurrent software
verification. In contrast, state-of-the-art tools for coverability
proved only 61, 51, or 33 of these 115 cases!
Moreover, and possibly due to the characteristics of the
application domain, even the simplest version of the technique ---based on the marking equation--- is
successful in 84 cases. 
As a second contribution, and inspired
by work on interpolation, we show that a dual version of the classical
set of constraints, equivalent in expressive power, can be used not
only to check safety, but to produce an inductive invariant.
While some existing solvers based on state-space
exploration can also produce such invariants, we show that inductive invariants obtained
through our technique are usually orders
of magnitude smaller.  Additionally, while we can use the 
SMT solver iteratively to minimize the invariant, the tool almost always
provides a minimal one at the first attempt.

% The paper is organized as follows.
% We introduce the basic notation in Section~\ref{sec_preliminaries}.
% We recall the marking equation and traps approach in
% Sections~\ref{sec_method_safety}
% and~\ref{sec_method_safety_by_refinement}.
% We introduce three novel methods to constructing invariants in
% Sections~\ref{sec_method_invariant}
% % \ref{sec_method_invariant_by_refinement},
% and
% \ref{sec_method_invariant_w_minimization}.
% Section~\ref{sec_experiments} presents (1) our experimental evaluation of
% the application of SMT solver \zthree\ to the constraint-based
% approach, and (2) our experimental evaluation of our methods for
% invariant construction.
% Section~\ref{sec_rel_work} reviews related
% work. Section~\ref{sec_conclusions} presents our conclusions.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: