### 2014-03-17 paper/workspace/sec_rel_work.tex

\smallskip
\noindent\textbf{Related Work.\ }
%
Our starting point was the work of Esparza and Melzer on
extending the marking equation with trap conditions to gain a stronger method
for proving safety of Petri nets \cite{EsparzaM00}. We combined the
constraint-based approach there with modern SMT solvers.
Their focus on (integer) linear programming tools of the time enforced some limitations. 
First, while traps are naturally encoded using Boolean variables,
\cite{EsparzaM00} encoded traps and the marking equation together into a set of 
linear constraints.
This encoding came at a practical cost: the encoding required (roughly) $n\times m$ constraints
for a Petri net with $n$ places and $m$ transitions, whereas the natural Boolean encoding
requires $m$ constraints.
Moreover, (I)LP solvers were not effective in searching large Boolean state
spaces; our use of modern SAT techniques alleviates this problem. 
%    constraints is cumbersome: it requires $n\times k$ constraints, where $n$ is
%    the number of places, and $k$ is the number of arcs from places to
%    transitions (succintly: $\#\{(p,t)\mid F(p,t)>0\}$. Expressing the trap
%    condition using boolean constraints requires only $m$ constraints, where $m$
%    is the number of transitions. 
Second, (I)LP solvers used by \cite{EsparzaM00} did not handle strict inequalities. 
Because of this, the authors used additional tricks, such as posing the problem that includes a 
strict inequality as a minimization problem, with the goal of minimizing the involved expression, 
and testing if the minimal value equaled zero. 
Unfortunately, this trick led to numerical instabilities. 
All of these concerns vanish by using an SMT solver.

In contrast to other recent techniques for coverability \cite{GantyFI08,KaiserKWCONCUR12,KloosMNP13},
our technique is incomplete.
However, we empirically demonstrate that for safe instances arising out of software verification,
our technique is extremely effective, often beating well-optimized techniques in terms of time or memory usage.
At the same time, our technique applies not only to coverability but also to \emph{reachability}.
It will be interesting to see whether the techniques can effectively verify liveness properties for concurrent
programs, which can be reduced to Petri net reachability questions.

% Main differences:
% \begin{itemize}
%     \item We have extended the techniques to construct inductive invariants,
%     which serve as certificates of safety.
%     \item Their focus on (integer) linear programming tools of the time enforced
%     some limitations. First, expressing the trap condition using linear
%     constraints is cumbersome: it requires $n\times k$ constraints, where $n$ is
%     the number of places, and $k$ is the number of arcs from places to
%     transitions (succintly: $\#\{(p,t)\mid F(p,t)>0\}$. Expressing the trap
%     condition using boolean constraints requires only $m$ constraints, where $m$
%     is the number of transitions. Second, (I)LP solvers of the time could not
%     handle strict inequalities. Because of this, tricks were used, such as
%     posing the problem that includes a strict inequality as a minimization
%     problem, with the goal of minimizing the involved expression, and testing
%     if the minimal value equals zero. Subsequently, this trick led to numerical
%     problems. All of these concerns vanish by using an SMT solver.
%     \item They focused their evaluation on proving safety mostly for mutual
%     exclusion protocols (check!). We have evaluated the methods on a
%     variety of examples that appear in software verification, provenance
%     analysis\ldots
% \end{itemize}
% 
% Other methods and tools \cite{KloosMNP13, KaiserKWCONCUR12, GeeraertsRB06,
% GantyFI08, ValmariH12}\ldots Even though the methods are complete, the tools
% suffer from time and memory limitations. However, they can also give refutations in the form of
% conterexamples.
% 
% Theoretical results \cite{KarpM69, Rackoff78}?